---
title: "a1_task4_lindsay_hafen.Rmd"
author: "Lindsay Hafen"
date: "10/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

## 1. View the data

- Attach the tidyverse and palmerpenguins in the initial code chunk library
- Use View funtion (Capital V) in **console**

## 2. Create a table with species and year

- Use dplyr::count to parse out columns species and year

```{r}
penguins %>% 
  count(species, year)
```

## 3. Create a scatterplot of flipper length vs bill depth that includes each species

- Select species, flipper length, and bill depth using dplyr::select function

```{r}
peng_filtered <- penguins %>% 
  select(species, bill_depth_mm, flipper_length_mm)
```

## 4. Create a graph of bill depth vs flipper length in respect to species

- Use ggplot to create a graph
- Assign variables x = flipper length, y = bill depth

```{r}
ggplot(data = peng_filtered, aes(x = flipper_length_mm, y = bill_depth_mm, color = species))+
  geom_point()+
  labs(title = "Flipper Length vs Bill Depth for Palmer's Penguins (2007-2009) (HAFEN)")+
  labs(x = "flipper length (mm)")+
  labs(y = "bill depth (mm)")+
  theme_dark()
```

## 5. Explore only gentoo 2008 data for flipper length vs body mass

- Create a new data subset peng_gent
- Use dplyr::select to parse out species, flipper length, body mass
- Use dplyr::filter to select year = 2008, species = gentoo

```{r}
peng_gent <- penguins %>% 
  select(species, flipper_length_mm, body_mass_g, year, sex) %>% 
  filter(species == "Gentoo") %>% 
  filter(year == 2008)
```

## 6. Graph flipper length vs body mass respective to sex for gentoo species in 2008

- Use ggplot to create a scatterplot using peng_gent for data 
- aes(x = flipper length, y = body mass, and color = sex)

```{r}
ggplot(data = peng_gent, aes(x = flipper_length_mm, y = body_mass_g, color = sex))+
  geom_point()+
  labs(title = "Flipper Length vs Body Mass Respective to Sex in Gentoo Palmer's Penguins (2008) (HAFEN)")+
  labs(x = "flipper length (mm)")+
  labs(y = "body mass (g)")+
  theme_classic()+
  theme(plot.title = element_text(size=11))
```






