---
title: "a1_task3_lindsay_hafen.Rmd"
author: "Lindsay Hafen"
date: "10/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
```

## 1. Read in Atlantic Salmon Imports data 

- Here I attach tidyverse in initial code chunk
- Read data using read function

```{r}
atl_salmon_imp_data <- read_csv("atl_salmon_imports.csv")
```

## 2. Change column titles to lower_snake_case

- I attach janitor in the initial code chunk
- I change the names using the clean_names function

```{r}
sal_clean <- atl_salmon_imp_data %>% 
  clean_names()
```

## 3. Wrangle data

- Here I will isolate variables: Salmon imports from Ireland, Denmark, and Iceland from 2000 - to 2018
- Use dplyr::filter()
- Is there a more concise way to filter each of these?

```{r}
sal_00to18 <- sal_clean %>% 
  filter(country_name == "Ireland" | country_name == "Denmark"| country_name == "Iceland") %>% 
  filter(import_year >= 2000) %>% 
  filter(import_year <= 2018)
```

## 4. Create a graph with 3 lines to signify the 3 respective countries

- Here I will create a graph with Ireland, Denmark and Iceland's salmon import by volume.
- I will use ggplot and modify each line independently with aes(color = country_name) (Is there a way to do this and select seperate colors for each country?)
- Why is aes(color = ) the argument used to change titles? Does it do something for colors?


```{r}
ggplot(data = sal_00to18, aes(x = import_year, y = salmon_volume, color = country_name)) +
  geom_line()+
  labs(title = "Estimated Volume of USA Salmon Imports (2000-2018) HAFEN")+
  labs(x = "year") +
  labs(y = "volume (10,000's of lbs)")+
  labs(color = "country of origin")
```



