---
title: "a1_task2_lindsay_hafen.Rmd"
author: "Lindsay Hafen"
date: "10/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
```

## 1. Read in St. Louis blood data

- Attach tidyverse using library(tidyverse)
- Read in data using read.csv

```{r}
stl_data <- read.csv("stl_lead.csv")
```

## 2. Clean St. Louis blood data

- Attach janitor using library(janitor)
- Clean data using clean_names() funtion

```{r}
stl_clean <- stl_data %>%  
  clean_names()
```

## 3. Wrangling the "clean" St. Louis blood data

- Using dplyr::
- select() for the percentage of children whose blood lead levels >= 5 micrograms/dL for: total population and black populations
- mutate() to create column: "Percent identifying as black"

```{r}
stl_2015 <- stl_clean %>% 
  select(pct_elevated, total_pop, black) %>% 
  mutate(perc_black = 100 * black / total_pop)
```

## 4. Create graph using new "% black population" variable against "% of children in each census tract with elevated blood lead levels"

- Change the color and size of the points in the geom_point() function
- Label the axes and title by adding labs()
- Change the size of the title by adding theme(plot.title = element_text(size = "a number")) (Thank you Google!) (are there other/preferred ways to accomplish this?)
- Change the theme of the graph using theme_"pick an option"

```{r}
ggplot(data = stl_2015, aes(x = perc_black, y = pct_elevated)) +
  geom_point(color = "red", size = .5) + labs(title = "Elevated Blood-Lead Levels in the Young Black Population of St. Louis, MO in 2015 (HAFEN)") +
  labs(x = "% of children identifying as black") +
  labs(y = "% of children with blood-lead levels >5 mcg/dL")+
  theme(plot.title = element_text(size=10)) + 
  theme_classic() 
```

## 5. Graph description

- This graph shows a general trend in which the percent of people in an area with high blood-lead levels increases as the percent of those people who identify as black increase. 




