---
title: "plastic_rivers_lindsay_hafen.Rmd"
author: "Lindsay Hafen"
date: "10/22/2020"
output: html_document
---

```{r setup, echo=TRUE, error=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
```

## 1. (a) Overview of exploration

This is an exploration of plastic pollution that results from macro and microplastic transported by rivers around the world. In this exploration I compare the top 15 receiving seas that are impacted by microplastic pollution, and I compare the top 5 contributing rivers' total plastic contribution to the Gulf of Mexico. I will practice using functions such as summarize and mutate as well as data visualization techniques that optimize responsible communication. 

#####(b) **Citation:**
Schmidt C, Krauth T, Wagner S (2017). Export of plastic debris by rivers into the sea. Environ. Sci. Technol. v. 51 (21), p. 12246-12253. 


## 2. (d) Read in the data & clean names

- Read in data using here() because it is in the data folder within the R Project
- Use janitor to change column names to lower_snake_case

```{r}
plastic_rivers <- read_csv(here("data", "plastic_rivers.csv" )) %>% 
  clean_names()
```

## 3. Clean up data

- Rename confusing columns using rename() function
- I'm taking units out here for brevity and readability, but they will be reflected in the graph later.
- MMPW stands for mismanaged plastic waste

```{r}
plastic_rivers_clean <- plastic_rivers %>% 
  rename(mmpw_generation_per_capita = mmpw_generation_per_capita_kg_d_1,
         microplastic_load_model_1 = microplastic_load_model_1_tons_y_1,
         mmpw_catchment_generated = mmpw_generated_in_the_catchment_tons_y_1,
         microplastic_load_model_2 = microplastic_load_model_2_tons_y_1,
         macroplastic_load = macroplastic_load_tons_y_1)
```

## 4. Explore data

- Use Model 2 to find total estimated annual microplastics for each receiving sea (Still in tonnes)
- Group data by receiving sea (combine all numeric values to correspond with their receiving sea)
- Use summarize to calculate total microplastics load across river catchments (that lead into receiving seas)
- Isolate the top 15 most impacted seas
- I will leave the East China Sea and the Yellow Sea separate because it seems that the original researchers may have recorded their data this way intentionally.

```{r, message=FALSE}
annual_microplastics <- plastic_rivers_clean %>% 
  select(microplastic_load_model_2, receiving_sea) %>%
  group_by(receiving_sea) %>% 
  summarize(total_microplastics = sum(microplastic_load_model_2, na.rm = TRUE))

top_15_impacted_seas <- annual_microplastics %>% 
  slice_max(order_by = total_microplastics, n = 15)
  
```

## 5. Graph of the top 15 impacted seas (microplastic tonnes/yr)

- I will use a bar graph to compare the impacted sea and the amount of microplastic (in tonnes)
- The data will be arranged in descending order with the most impacted sea at the top.
- I have elected to place the sea names on the y axis for readability, but otherwise a y axis title is unnecessary. I have removed the y axis title using the theme() function

```{r}
ggplot(top_15_impacted_seas, 
       aes(y = fct_reorder(receiving_sea, total_microplastics),
           x = total_microplastics)) +
  geom_bar(stat = "identity", fill = "darkcyan") +
  labs(title = "Top 15 seas annually impacted by microplastics", x = "microplastics (tonnes per year)") +
  theme_minimal() +
  theme(axis.title.y = element_blank())
```

##### *fig(1)* 
graph depicts top 15 seas impacted by microplastic particles (each <5mm) annually measured in estimated tonnes of plastic.

## 6. Graph of the top 5 catchments for plastic pollution in the Gulf of Mexico

- First I will need to organize the data starting from our cleaned plastics
- dplyr::select to select for model 2 micro and macroplastics, rivers and receiving sea
- dplyr::filter to select only catchments that have a receiving sea of the Gulf of Mexico (GoM)
- Combine the micro and macroplastics values into one column
- I'm removing the macroplastic_load and microplastic_load_model_2 columns for practice
- use slice_max() with an n = 5 to isolate the top 5 catchment contributors to the GoM
- Graph a bar graph that reflects the river catchments vs the tonnage of plastic into the gulf per year.

```{r}
top_5_catchments <- plastic_rivers_clean %>% 
  select(river, receiving_sea, microplastic_load_model_2, macroplastic_load) %>% 
  filter(receiving_sea == "Gulf of Mexico") %>% 
  mutate(total_plastics = microplastic_load_model_2 + macroplastic_load) %>% 
  select(- microplastic_load_model_2, - macroplastic_load) %>% 
  slice_max(order_by = total_plastics, n = 5)

ggplot(data = top_5_catchments, 
       aes(y = fct_reorder(river, total_plastics), x = total_plastics)) +
  geom_bar(stat = "identity", fill = "darkmagenta") +
  labs(title = "Top five annual contributers of plastic to the Gulf of Mexico", x = "tonnes of plastic") +
  theme_bw() +
  theme(axis.title.y = element_blank())
```

##### *fig(2)* 
This graph depicts the estimated tonnes of plastic that each of these top 5 contributor catchments contributes to the Gulf of Mexico annually.

- The top 5 contributers to the greatest total plastic loads in the Gulf of Mexico are:
1) Mississippi
2) Grijalva, Usumacinta
3) Rio Grande
4) Tamesi, Panuco
5) Trinity


